\normalfont{A bird's-eye view on the habitability of exoplanets via statistical learning techniques}

### 2020-10-03 Manuscript/Manuscript.tex
PUB

\subparagraph{Linear discriminat analysis}
Here the density $f(x)$ is assumed to be a multivariate gaussian distribution\cite{james2013introduction}: 
\begin{equation}
f(\textbf{x})=\dfrac{1}{\left(2 \pi\right)^{p/2}\lvert \boldsymbol {\Sigma}\rvert^{-1/2} }exp \left(-\frac{1}{2}\left(\textbf{x}-\boldsymbol{\mu}\right)^{T}\boldsymbol {\Sigma}^{-1}\left(\textbf{x}-\boldsymbol{\mu}\right)\right)
\end{equation}
While the covariance matrix $\boldsymbol {\Sigma}$ is common for all classes while $\boldsymbol{\mu}$ is different for each of them. Thus, by some manipulation one obtains that the Bayes classifier put an observation $X=\textbf{x}$ for which \cite{james2013introduction}:
\begin{equation}
\delta(\textbf{x})=x^{T} \boldsymbol {\Sigma}^{-1} \mu_{k}-\frac{1}{2}\mu^{T}_{k}\boldsymbol {\Sigma}^{-1}\mu_{k}+log\pi_{k}
\end{equation}
is largest. Note the linearity of the expression with respect to the \textbf{x} vector, that give the name to the method 
\subparagraph{Quadratic discriminat analysis}
In this case, differently with respect to the linear analysis, a covariance matrix for each class is considered: thus the Bayes classifier has the following form 
\begin{equation}
\delta(\textbf{x})=-\dfrac{1}{2}\textbf{x}^{T}\boldsymbol {\Sigma}^{-1}_{k}\textbf{x}+\textbf{x}^{T}\boldsymbol {\Sigma}^{-1}_{k}\boldsymbol{\mu}_{k}-\dfrac{1}{2}\boldsymbol{\mu}_{k}^{T}\boldsymbol {\Sigma}^{-1}_{k}\boldsymbol{\mu}_{k}-\dfrac{1}{2}log\lvert \boldsymbol {\Sigma}\rvert+log\pi_{k}
\end{equation}
where in this case the dependence from $x$ is quadratic. Since QDA has a covariance matrix for each class, it requires a larger parameter estimation, thus a larger dataset is required for it with respect to the LDA. On the other side the assumption of common covariance matrix expose the LDA to the bias error.

@@

\begin{table}[]
\caption{Performances of LDA algorithm for the exoplanets habitability, evaluated by the accuracy, the no-information rate (NIR), and the $\phi$ factor.}
\begin{center}
\begin{tabular}{l|l}
\multicolumn{1}{c|}{Feature} & Value \\ \hline
\multicolumn{1}{c|}{Accuracy} & 0.957 \\ \hline
\multicolumn{1}{c|}{NIR} & 0.898 \\ \hline
\multicolumn{1}{c|}{$\phi$} & 0.738 \\ 
\label{LDA_perf_tab}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Performances of QDA algorithm for the exoplanets habitability, evaluated by the accuracy, the no-information rate (NIR), and the $\phi$ factor.}
\begin{center}
\begin{tabular}{l|l}
\multicolumn{1}{c|}{Feature} & Value \\ \hline
\multicolumn{1}{c|}{Accuracy} & 0.957 \\ \hline
\multicolumn{1}{c|}{NIR} & 0.899 \\ \hline
\multicolumn{1}{c|}{$\phi$} & 0.887 \\ 
\label{QDA_perf_tab}
\end{tabular}
\end{center}
\end{table}

@@

>>>>

\subparagraph{Linear discriminat analysis}
Here the density $f(x)$ is assumed to be a multivariate gaussian distribution\cite{james2013introduction}: 
\begin{equation}
f(\textbf{x})=\dfrac{1}{\left(2 \pi\right)^{p/2}\lvert \boldsymbol {\Sigma}\rvert^{-1/2} }exp \left(-\frac{1}{2}\left(\textbf{x}-\boldsymbol{\mu}\right)^{T}\boldsymbol {\Sigma}^{-1}\left(\textbf{x}-\boldsymbol{\mu}\right)\right)
\end{equation}
While the covariance matrix $\boldsymbol {\Sigma}$ is common for all classes while $\boldsymbol{\mu}$ is different for each of them. Thus, by some manipulation one obtains that the Bayes classifier put an observation $X=\textbf{x}$ for which \cite{james2013introduction}:
\begin{equation}
\delta(\textbf{x})=x^{T} \boldsymbol {\Sigma}^{-1} \mu_{k}-\frac{1}{2}\mu^{T}_{k}\boldsymbol {\Sigma}^{-1}\mu_{k}+log\pi_{k}
\end{equation}
is largest. Note the linearity of the expression with respect to the \textbf{x} vector, that give the name to the method 
\subparagraph{Quadratic discriminat analysis}
In this case, differently with respect to the linear analysis, a covariance matrix for each class is considered: thus the Bayes classifier has the following form \cite{james2013introduction}:  
\begin{equation}
\delta(\textbf{x})=-\dfrac{1}{2}\textbf{x}^{T}\boldsymbol {\Sigma}^{-1}_{k}\textbf{x}+\textbf{x}^{T}\boldsymbol {\Sigma}^{-1}_{k}\boldsymbol{\mu}_{k}-\dfrac{1}{2}\boldsymbol{\mu}_{k}^{T}\boldsymbol {\Sigma}^{-1}_{k}\boldsymbol{\mu}_{k}-\dfrac{1}{2}log\lvert \boldsymbol {\Sigma}\rvert+log\pi_{k}
\end{equation}
where in this case the dependence from $x$ is quadratic. Since QDA has a covariance matrix for each class, it requires a larger parameter estimation, thus a larger dataset is required for it with respect to the LDA. On the other side the assumption of common covariance matrix expose the LDA to the bias error.

@@

\begin{table}[]
\caption{Performances of multiple attribute logistic regression algorithm for the exoplanets habitability, evaluated by the accuracy, the no-information rate (NIR), and the $\phi$ factor.}
\begin{center}
\begin{tabular}{l|l}
\multicolumn{1}{c|}{Feature} & Value \\ \hline
\multicolumn{1}{c|}{Accuracy} & 0.914 \\ \hline
\multicolumn{1}{c|}{NIR} & 0.869 \\ \hline
\multicolumn{1}{c|}{$\phi$} & 0.566 \\ 
\label{LDA_perf_tab}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Performances of LDA algorithm for the exoplanets habitability, evaluated by the accuracy, the no-information rate (NIR), and the $\phi$ factor.}
\begin{center}
\begin{tabular}{l|l}
\multicolumn{1}{c|}{Feature} & Value \\ \hline
\multicolumn{1}{c|}{Accuracy} & 0.902 \\ \hline
\multicolumn{1}{c|}{NIR} & 0.898 \\ \hline
\multicolumn{1}{c|}{$\phi$} & 0.526 \\ 
\label{LDA_perf_tab}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Performances of QDA algorithm for the exoplanets habitability, evaluated by the accuracy, the no-information rate (NIR), and the $\phi$ factor.}
\begin{center}
\begin{tabular}{l|l}
\multicolumn{1}{c|}{Feature} & Value \\ \hline
\multicolumn{1}{c|}{Accuracy} & 0.941 \\ \hline
\multicolumn{1}{c|}{NIR} & 0.869 \\ \hline
\multicolumn{1}{c|}{$\phi$} & 0.757 \\ 
\label{QDA_perf_tab}
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}[h]
  \centering
\subfloat[]{\includegraphics[width=0.6\textwidth]{Pic/LOG_ROC.png}}\\
\subfloat[]{\includegraphics[width=0.6\textwidth]{Pic/LOG_FULL.png}}
\caption{The performances of the multiple logistic regression on the habitability of exoplanets (as obtained by a 10-fold cross validation) once the the dimensionality reduction via was applied (to 2 dimension): on the panel (a) the confusion matrix is given  while on panel (b) the ROC curve (lower panel) is reported. A predicted (P) habitable exoplanet is classified as 1 while a test (T) non-habitable exoplanet is classified as 0.}
\label{SVM+PCA_results}
\end{figure}
